---
title: "Analyze Shoes Data"
author: "Samir Gadkari"
date: "2/20/2021"
output: html_document
---

Data Source: [Shoes dataset](https://datafiniti.co)

```{r}
library(tidyverse)
library(lubridate)
```

# 1. Import data
```{r}
shoes <- read_csv("../mens_shoe_prices.csv")
```
There are 29 parsing problems. Let's see if the rest of our data
looks good.

```{r}
nrow(shoes)
ncol(shoes)
head(shoes)
```

We have a lot of rows (other than the 29 parsing failures).
Let's continue with what we have.

# 2. Analyze
## 2.1 Select columns and rows

Let's look at which columns have many NA values.
Then we can decide which ones to use.
```{r}
fraction_not_missing <- function(x) {
  mean(!is.na(x))
}

x <- sort(unlist(map(shoes, fraction_not_missing)), 
          decreasing = TRUE)

# Putting the values in a dataframe just makes it easier to read
df <- tibble(name = names(x), values = as.vector(x))
knitr::kable(df)
```

Some things to note about this dataframe:

  * There are many other items than shoes.
  * The id is not unique - we may not use it, but it's good to note this
  * categories contain values from the features column's JSON string
  * dateadded, dateupdated: We will need these if we're looking at timelines.
                            Are there many rows with different values between these
                            columns?
  * keys: contain the brand name / manufacturer number. These have already been parsed
          into brand and manufacturernumber columns.
          The brand column is almost fully populated.
          78% of the manufacturernumber column is populated
  * name: What you would expect to see as the name of the shoe
  * prices_amountmax
  * prices_amountmin
  * prices_dateadded
  * prices_dateseen: What is the difference between prices_dateadded and prices_dateseen?
  * prices_issale
  * prices_currency: What are all the currencies in this dataset? 
                     If only one currency, we can ignore this column.
  * prices_sourceurls: Contains the Website and merchant names.
                       This has been extracted into the prices_merchant column,
                       so we can drop this column.
  * sourceurls: Looks the same as prices_sourceurls. Drop it.
  * brand: We can use this to compare prices by brand
  * imageurls: We don't need this. Even if we did, there is no guarantee that
               the images exist anymore on the website. Drop it.
  * manufacturernumber: We can use this to see what price range different manufacturers 
                        target
  * merchants: Contains "dateSeen" and "name"
  * features: Provides color, dimensions, etc.
  * prices_merchant: Contains seller name and website where product is sold
  
### 2.1.1 Get columns with only mens shoes

Here we decided not to get sandals, slippers, or crocs since they're
not shoes. 

We tried using the categories column, but it labeled
"Mens Shoes" for sandals, infant shoes, and shoe studs.
You can check the names column
```{r}
head((shoes %>% 
  filter(str_detect(categories, "[Mm]en")) %>%
  filter(str_detect(categories, "[Ss]hoe")))["name"], n = 15)
```

This is more precise:

```{r}
mens_shoes <- shoes %>%
  filter(str_detect(name, "[Mm]en"))
mens_shoes <- mens_shoes %>%
  filter(str_detect(name, "[Ss]hoe") | 
         str_detect(name, "[Bb]oot") |
         str_detect(name, "[Ll]oafer") | 
         str_detect(name, "[Ss]tep") |
         str_detect(name, "[Ss]neaker") | 
         str_detect(name, "[Tt]rainer") |
         str_detect(name, "[Oo]xford") | 
         str_detect(name, "[Cc]anvas"))
nrow(mens_shoes)/nrow(shoes)
head(mens_shoes["name"], n = 15)
```

That gives us a much smaller dataframe - it contains only 34% 
of the original number of rows.

### 2.1.2 Are dateadded, dateupdated, and prices_dateseen different?

```{r}
dateadded <- mens_shoes$dateadded
dateupdated <- mens_shoes$dateupdated
prices_dateseen <- mens_shoes$prices_dateseen
typeof(dateadded)
typeof(dateupdated)
typeof(prices_dateseen) # This is a character vector -  convert it.
prices_dateseen <- ymd_hms(prices_dateseen)

mean(!(dateadded == dateupdated), na.rm = TRUE)
mean(!(dateupdated == prices_dateseen), na.rm = TRUE)
mean(!(prices_dateseen == dateadded), na.rm = TRUE)
```

So we see that dateadded and dateupdated are different 1/3 of the time.
dateupdated and prices_dateseen are 98% different.
prices_dateseen and dateadded are 94% different.

Let's see if they're different by more than a day:
```{r}
mean(abs(dateadded - dateupdated) > days(1), na.rm = TRUE)
mean(abs(dateupdated - prices_dateseen) > days(1), na.rm = TRUE)
mean(abs(prices_dateseen - dateadded) > days(1), na.rm = TRUE)
```

This output is similar to the output earlier with:

  * 31% difference of more than a day between dateadded and dateupdated
  * 87% difference of more than a day between dateupdated and 
    prices_dateseen
  * 84% difference of more than a day between prices_dateseen and
    dateadded

This means there is information in these dates that might be useful.
According to the schema, dateadded was the date when the original
item was added. dateupdated was the date of the last product update.
prices_dateseen was the date that the given price was last seen.

### 1.2.3 Can we restrict ourselves to just one currency?

```{r}
mens_shoes %>% count(prices_currency, sort = TRUE)
```

Looks like USD is mostly use - let's remove all other rows and
the currency column.

```{r}
nrow(mens_shoes)
mens_shoes <- mens_shoes %>% 
  filter(prices_currency == "USD") %>%
  select(-prices_currency)
nrow(mens_shoes)
```

That makes sense - we lost around 170 rows. This is around what
we see when we sum up the frequencies of non-USD currencies above.

### 1.2.4 Drop all columns we don't need

```{r}
ncol(mens_shoes)
mens_shoes <- mens_shoes %>%
  select(-prices_condition, -colors, -upc, -ean, -descriptions,
         -skus, -manufacturer, -prices_offer, -sizes,
         -prices_shipping, -dimension, -asins, -reviews,
         -prices_returnpolicy, -weight, -prices_color,
         -prices_size, -prices_availability, -prices_warranty,
         -prices_count, -prices_flavor, -vin, -quantities,
         -websiteids, -count, -flavors, -isbn, -prices_source,
         -prices_sourceurls, -sourceurls, -imageurls)
ncol(mens_shoes)
```

# 2. Analyse dataset

## 2.1 Which brands are cheaper/costlier